### 2018-03-02 Writing/disparity_review.tex
metrics

Disparity metrics are essentially a one dimensional summary of a multidimensional object, hence, they are bound to only reflect some aspect of the morphospace and can never encompass its whole complexity.
With that in mind, the choice of the metric should be essential to summarise the aspect of the morphospace that is the most related to the question.
In an unidimensional distributions (e.g. histograms) different metrics will readily reflect different aspect of the data.
For example, metrics can be used to reflect the ``size'' of the distribution (e.g. the range, quantiles or variance) and some other will reflect the most likely value (the central tendency - e.g. the mean, median or mode).
Among these metrics, some will have more preferable properties than other such as sensitivity to outliers (range, mean or mode \textit{vs.} quantiles, variance or median) and will thus be more or less appropriate to the different questions.

In multidimensionality, this high availability of unidimensional descriptors of the space are really numerous and come with their equally numerous advantages and caveats.
In this case, the disparity question should again drive the choice of the disparity metric.
If the question refers to ``size'' of a group in the morphospace (e.g. ``does group A occupy as much space as group B?''), metrics relating to the spread of the group in the morphospace are the most appropriated (e.g. volume [CITE], variance [CITE], range [CITE], distance from the centroid [CITE]).
Conversely, if the question refers to the ``position'' of a group in the morphospace (e.g. ``does group A occupy the same space as group B?''), metrics relating to the distance of the elements to a point are most appropriated (e.g. distance from a fixed point - e.g. the opposite group).
Finally, one might also be interested in the density of elements in some regions of the morphospace (e.g. ``is group A more dense than group B?''), metrics relating to the pairwise distances between elements are most appropriated (e.g. nearest neighbour distance, pairwise distances, etc.)

>>>>

After building and visualising the morphospace, one can use disparity metrics (or indices) as a one dimensional summary of it to test hypotheses.
Because of these reduction in dimensionality, these metrics are bound to only reflect some aspect of the morphospace and can never encompass its whole complexity.
With that in mind, the choice of the metric should be essential to summarise the aspect of the morphospace that is the most related to the question.
In an unidimensional distributions (e.g. histograms) different metrics will readily reflect different aspect of the data.
For example, metrics can be used to reflect the ``size'' of the distribution (e.g. the range, quantiles or variance) and some other will reflect the most likely value (the central tendency - e.g. the mean, median or mode).
Among these metrics, some will have more preferable properties than other such as sensitivity to outliers (range, mean or mode \textit{vs.} quantiles, variance or median) and will thus be more or less appropriate to the different questions.

In multidimensionality, this high availability of unidimensional descriptors comes with equally numerous advantages and caveats.
If the question refers to ``size'' of a group in the morphospace (e.g. ``does group A occupy as much space as group B?''), metrics relating to the spread of the group in the morphospace are the most appropriated (e.g. volume [CITE], variance [CITE], range [CITE], distance from the centroid [CITE]).
Conversely, if the question refers to the ``position'' of a group in the morphospace (e.g. ``does group A occupy the same space as group B?''), metrics relating to the distance of the elements to a point are most appropriated (e.g. distance from a fixed point - e.g. the opposite group).
Finally, one might also be interested in the density of elements in some regions of the morphospace (e.g. ``is group A more dense than group B?''), metrics relating to the pairwise distances between elements are most appropriated (e.g. nearest neighbour distance, pairwise distances, etc.)
